### 2021-07-05 rad-icfp-2021/intro.tex
Listings improvements + Fixed acmart error

Our contribution makes these claims precise. In more details, we introduce
\begin{itemize}
    \item an expressive source language with array operations, yet restrictive enough for efficient differentiation (\S\ref{sub:sourcelang})
    \item a method for extending our work and modular approaches to add and prove correct new primitives as needed in practice (\S\ref{sec:generalization})
    \item a new intermediate representation (IR), UNF, essentially turning every operation into a unary one which inputs and outputs a tuple (\S\ref{sec:unf})
    \item a transformation between the IR and the main language (\S\ref{sub:transformations to and from UNF})
    \item a simple reverse-mode transformation on the IR (\S\ref{sub:Simple reverse mode transformation})
    \item the direct reverse-mode transformation obtained from the source to the target language (\S\ref{sub:Macro for pure reverse mode transformation})
    \item denotational semantics of our languages and transformations using multicategories and concategories (\S\ref{sec:correctness})
    \item correctness of the transformations (\S\ref{sec:correctness})
    \item complexity guarantees (\S\ref{sec:complexity})
    \item how efficiency is recovered via standard optimization techniques (\S\ref{sub:Optimizations})
    \item several additions to the source language (\S\ref{sec:generalization}) \MH{should emphasize that it should be easier for us to add primitives compared to several recent work. give recipe.}
    \item a detailed comparison with other work and some limitations to our work (\S\ref{sec:conclusion})
\end{itemize}

@@

We introduce the general idea of how we do efficient reverse-mode in a functional setting on some examples.
We recall some basics of reverse-mode differentiation in Section~\ref{sec:background}.

\begin{example}[First-order term]
On simple first-order terms, the output code looks like SSA code \cite{cytron1989efficient}.
Let us consider  the term !let w1 = x1 * x2 in let w2 = w1 * x1 in w2! in the context $\Gamma:=\{x1,x2,x3:\RR\}$.\\
After our (inefficient) reverse-mode transformation, we obtain
\begin{center}
    \begin{tabular}{l}
        !let w1,w1'! = !< x1*x2, fun (y1,...,y4) -> (y1+x2*y4, y2+x1*y4, y3)> in!\\
        !let w2,w2'! = !< w1*x1, fun (y1,...,y5) -> w1'!!(y1+w1*y5, y2, y3, y4+x1*y5)> in!\\
        !w_2'(0,0,0,0,1)!
    \end{tabular}
\end{center}
The part !(0,0,0,0,1)! comes as an equivalent to initialing the tangent variables in the imperative reverse-mode algorithm.

>>>>

Our contribution makes these claims precise. In more details, we introduce
\begin{itemize}
    \item an expressive source language with array operations, yet restrictive enough for efficient differentiation (\S\ref{sub:sourcelang})
    \item a method for extending our work and modular approaches to add and prove correct new primitives as needed in practice (\S\ref{sec:generalization})
    \item a new intermediate representation (IR), UNF, essentially turning every operation into a unary one which inputs and outputs a tuple (\S\ref{sec:unf})
    \item a transformation between the IR and the main language (\S\ref{sub:transformations to and from UNF})
    \item a simple reverse-mode transformation on the IR (\S\ref{sub:Simple reverse mode transformation})
    \item the direct reverse-mode transformation obtained from the source to the target language (\S\ref{sub:Macro for pure reverse mode transformation})
    \item a denotational semantics of our languages and transformations using multicategories and concategories (\S\ref{sec:correctness})
    \item correctness of the transformations (\S\ref{sec:correctness})
    \item complexity guarantees (\S\ref{sec:complexity})
    \item how efficiency is ensured via standard optimization techniques (\S\ref{sub:Optimizations})
    \item several additions to the source language (\S\ref{sec:generalization}) \MH{should emphasize that it should be easier for us to add primitives compared to several recent work. give recipe.}
    \item a detailed comparison with other work and some limitations of our work (\S\ref{sec:conclusion}) \YK{Is it really necessary to specify this point in the contributions? (as it is supposed to be part of every paper)}
\end{itemize}

@@

We introduce the general idea of efficient reverse-mode in a functional setting on some examples.
We recall some basics of reverse-mode differentiation in Section~\ref{sec:background}.

\begin{example}[First-order term]
On simple first-order terms, the output code is reminiscent of SSA code~\cite{cytron1989efficient}.
Let us consider the term !let w$_1$ = x$_1$ * x$_2$ in let w2 = w1 * x1 in w2! in the context $\Gamma:=\{x_1,x_2, x_3:\RR\}$.\\
After our (inefficient) reverse-mode transformation, we obtain
\begin{center}
    \begin{tabular}{l}
        !let w1,w1'! = !< x1*x2, fun (y1,...,y4) -> (y1+x2*y4, y2+x1*y4, y3)> in!\\
        !let w2,w2'! = !< w1*x1, fun (y1,...,y5) -> w1'!!(y1+w1*y5, y2, y3, y4+x1*y5)> in!\\
        !w_2'(0,0,0,0,1)!
    \end{tabular}
\end{center}
The part !(0,0,0,0,1)! comes as an equivalent to initialing the tangent variables in the imperative reverse-mode algorithm.
